---
title: "Yield_analysis"
output: html_document
---
```
Initial ML workflow drafted by David Makowski INRAE (18 Feb 2021)
Additional ML analyses, Shapley explanations, and Figures developed by Rose Abramoff (22 Feb 2021)

This analysis has two main sections:
1) testing the performance of different random forest models and generating plots
2) calculating Shapley additive explanations and generating plots

The analysis sources functions:
multiplot.R (for plotting)
readin.R (for reading in data)
plotfuncs.R (for plotting)
yieldloop.R (for testing ML models)
shaploop.R (for calculating Shapley values)
```
#Load libraries
```{r load-libraries}
rm(list=ls()) 

library(openxlsx)
library(dplyr)
library(ggplot2)
library(spaMM)
library(ranger)
library(gbm)
library(iml)
library(lme4)
library(randomForest)
library(mapdata)
library(raster)
library(DALEX)
library(caret)
#library(tidyverse)
```

#Switches
```{r switches}
do_maps = T
lowcolor = "#440154FF"
midcolor = "#21908CFF"
highcolor = "#FDE725FF"
  
lowcolor = "#533600"
midcolor = "#F6F6F6" 
highcolor = "#004B40" 
  
diverging_hcl(palette = "Green-Brown",n = 9)
```

#Data import
```{r data-import}
rootdir <- "/Users/rzabramoff/Documents/ORNL/DropboxORNL/Collaborations/ProjectYield"
outputdir= paste0(rootdir,"/Output/Rdata/")
figoutdir= paste0(rootdir,"/Output/Figures/")
crop_species = c("Maize","Rice","Wheat","Soybean")
source('multiplot.R')
source('readin.R')
source('plotfuncs.R')
source('yieldloop.R')
source('shaploop.R')
FTAB <- readin()
```

#Do Maps
```{r, do-maps}
if(do_maps){
pdf(file=paste0(figoutdir,"yield_meta.pdf"), height=8, width=10)
print(map_of_df(FTAB))
dev.off()

pdf(file=paste0(figoutdir,"yield_hist.pdf"), height=6, width=6)
print(hist_of_df(FTAB))
dev.off()
}
```

#Describe data
```{r}
FTAB %>% 
  group_by(Crop) %>% 
  dplyr::summarize(locations = length(unique(Loca)),
            samples = length(Effect),
            meanEffect = mean(Effect),
            medianEffect = median(Effect),
            minEffect = min(Effect),
            maxEffect = max(Effect),
            sd = sd(Effect),
            se = sd(Effect)/sqrt(length(Effect)))

shorter <- FTAB %>%
  dplyr::select(Effect,Crop,Delta_temp,CO2.ppm,TempAvg,Precip,Delta_precip,Latitude,Longitude,Fertiliser,Irrigation,Cultivar,Soil.organic.matter.management,Planting.time,Tillage,Others,Loca,Adaptation.type) %>%
  tidyr::drop_na()

length(unique(shorter$Loca))

table(shorter$Adaptation.type, by=shorter$Crop)
```

#Yield Loop
```{r, eval=F}
#For training
start <- Sys.time()
loop_func(same_locations = T, adaptation_types = F, do_iml = F, Nb = 10, run_all_mods = T, train_on_all = F)
end <- Sys.time()
end-start
#Approx 45 mins

start <- Sys.time()
loop_func(same_locations = F, adaptation_types = F, do_iml = F, Nb = 10, run_all_mods = T, train_on_all = F)
end <- Sys.time()
end-start
#Approx 30 mins

#For training
start <- Sys.time()
loop_func(same_locations = T, adaptation_types = F, do_iml = T, Nb = 10, run_all_mods = F, train_on_all = F)
end <- Sys.time()
end-start
#Approx 2.6 hr

start <- Sys.time()
loop_func(same_locations = F, adaptation_types = F, do_iml = T, Nb = 10, run_all_mods = F, train_on_all = F)
end <- Sys.time()
end-start

same_locations = F
adaptation_types = T
do_iml = T
Nb = 1
run_all_mods = F
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start

same_locations = F
adaptation_types = T
do_iml = F
Nb = 1
run_all_mods = F
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start

same_locations = F
adaptation_types = F
do_iml = F
Nb = 1
run_all_mods = F
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start

#Make 2 way interaction plot
same_locations = F
adaptation_types = F
do_iml = T
Nb = 1
run_all_mods = F
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start

#output lm summaries
same_locations = F
adaptation_types = F
do_iml = F
Nb = 1
run_all_mods = T
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start
```

##Make Plots
###Make obs v pred pred plots
```{r}
same_locations = T #run both T and F
adaptation_types = F
do_iml = F
Nb = 10
run_all_mods = T
train_on_all = F
fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_store_bestML_predictions.Rdata"))

pdf(file=paste0(figoutdir,"FigureS3_samelocs",same_locations,"_adaptypes",adaptation_types,".pdf"), height=4, width=6)
obs_v_pred(ObsPred)
dev.off()

resultdf <- read.csv(paste0(fileprefix,"_results.csv"))
resultdf

  AllOP <- bind_rows(ObsPred, .id = "column_label")
  FormOP <- AllOP %>%
    tidyr::pivot_longer(cols=c(Predicted, Observed), names_to="Type") %>%
    group_by(Crop, Site, Type) %>%
    
    dplyr::summarise(mean = mean(value),
                     se = sd(value)/sqrt(n()),
                     n = n()) %>%
    tidyr::pivot_wider(names_from = Type, names_sep = ".", values_from = c(mean, se, n)) %>% mutate(Residual=mean.Observed - mean.Predicted)
  
pdf(file=paste0(figoutdir,"Residuals_samelocs",same_locations,"_adaptypes",adaptation_types,".pdf"), height=4, width=6)
  plot(FormOP$mean.Predicted, FormOP$Residual, ylab="Residuals", xlab="Observed Yield Change")
dev.off()
```

###Make PDP plots
####Load the data
```{r}
same_locations = F
adaptation_types = T
do_iml = T
Nb = 1
run_all_mods = F
train_on_all = T
fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_store_bestML_PDPs.Rdata"))
#PDPeffs[[Nb 1-10]][[Crop 1-4]]
```

####Make the plots
```{r}
cbPalette <- c("#000000","#E69F00", "#009E73", "#0072B2", "#CC79A7", "#D55E00", "#56B4E9", "#F0E442")

bindeffs <- function(PDPeffs, Crop.Num) {
  effs <- NULL
for(i in 1:1){
effs[[i]] <- do.call("rbind", PDPeffs[[i]][[Crop.Num]]$results)
effs[[i]]$.crop <- rep(crop_species[Crop.Num], dim(effs[[i]])[1])
}
neweffs <- do.call("rbind", effs)
return(neweffs)
}

maize.effs <- bindeffs(PDPeffs, 1)
rice.effs <- bindeffs(PDPeffs, 2)
wheat.effs <- bindeffs(PDPeffs, 3)
soybean.effs <- bindeffs(PDPeffs, 4)
soybean.effs$.borders <- as.character(soybean.effs$.borders)

     efdf <- maize.effs %>% 
       dplyr::full_join(rice.effs) %>%
       dplyr::full_join(wheat.effs) %>%
       dplyr::full_join(soybean.effs) %>%
        group_by(.borders, .feature, .crop) 
     efdf$.crop <- as.factor(efdf$.crop)
     
if(adaptation_types){
    efdf.adaptation <- filter(efdf, .feature != "TempAvg" && .feature != "CO2.ppm" && .feature != "Delta_temp" && .feature != "Latitude" && .feature != "Longitude" && .feature != "Precip" && .feature != "Delta_precip")
    efdf.others <- filter(efdf, .feature == "TempAvg" | .feature == "CO2.ppm" | .feature == "Delta_temp" | .feature == "Latitude" | .feature == "Longitude" | .feature == "Precip" | .feature == "Delta_precip")
    
    efdf.adaptation$.feature <- as.factor(efdf.adaptation$.feature)
    levels(efdf.adaptation$.feature) <- c("Cultivar", "Fertilizer", "Irrigation", "Others", "Planting Time")
    efdf.adaptation <- efdf.adaptation %>%
      tidyr::pivot_wider(names_from=.borders, values_from=.value) %>%
      mutate(.value = Yes-No)
}else{
    efdf.adaptation <- efdf[efdf$.feature=="Adaptation",]
    efdf.others <- efdf[efdf$.feature!="Adaptation",]
}
     
     efdf.others$.borders <- as.numeric(efdf.others$.borders)
     efdf.TempAvg <- efdf.others[efdf.others$.feature=="TempAvg",]
     efdf.Delta_temp <- efdf.others[efdf.others$.feature=="Delta_temp",]
     efdf.CO2.ppm <- efdf.others[efdf.others$.feature=="CO2.ppm",]
     efdf.Latitude <- efdf.others[efdf.others$.feature=="Latitude",]
     efdf.Longitude <- efdf.others[efdf.others$.feature=="Longitude",]
     efdf.Precip <- efdf.others[efdf.others$.feature=="Precip",]
     efdf.Delta_precip <- efdf.others[efdf.others$.feature=="Delta_precip",]
    
      ylabs = "Yield Change (%)"
      ylims = range(efdf$.value)+range(efdf$.value)*0.1
      
      ggplot_pdps <- function(efdf, xlabels) {
      ggplot(efdf, aes(.borders, .value, color= .crop)) +
        geom_point(alpha=0.3) + geom_line() +
        ylim(ylims) +
        ylab(ylabs) +
        xlab(xlabels) + 
        theme_bw() + theme(legend.position = "none") +
        scale_color_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4])) + 
        guides(colour = guide_legend(override.aes = list(size=5)))
      }
      
      ggplot_pdps_latlong <- function(efdf, xlabels) {
      ggplot(efdf, aes(.borders, .value, color= .crop)) +
        geom_point(alpha=0.3) + geom_line() +
        ylim(ylims) +
        ylab(ylabs) +
        xlab(xlabels) + 
        theme_bw() + theme(legend.position = "none") +
        scale_color_manual(values = c(cbPalette[2], cbPalette[3], cbPalette[4])) + 
        guides(colour = guide_legend(override.aes = list(size=5)))
      }
      
p1 <- ggplot_pdps(efdf.TempAvg, expression("Average Temperature"~degree~"C"))
p2 <- ggplot_pdps(efdf.Delta_temp, expression(Delta~" Temperature"~degree~"C"))
p3 <- ggplot_pdps(efdf.CO2.ppm, expression(Delta~"CO"[2]~"ppm above 390 ppm"))
p4 <- ggplot_pdps_latlong(efdf.Latitude, "Latitude")
p5 <- ggplot_pdps_latlong(efdf.Longitude, "Longitude")
p8 <- ggplot_pdps(efdf.Precip, expression("Annual Precipitation (mm)"))
p9 <- ggplot_pdps(efdf.Delta_precip, expression("Annual "~Delta~" Precipitation (mm)"))

if(adaptation_types){
  p6 <- ggplot(efdf.adaptation, aes(x = .feature, y = .value, fill=.crop)) +
          geom_bar(stat = "identity", position="dodge") + 
          facet_grid(.crop~.) +
          ylab("Yield Improvement with Adaptation (%)") +
          xlab("") + 
          theme_bw() + theme(legend.position = "none", strip.text.x = element_text(size = 5), axis.text.x=element_text(angle = 45, hjust= 1)) +
          scale_fill_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[4])) + 
          guides(colour = guide_legend(override.aes = list(size=5)))
}else{
  p6 <- ggplot(efdf.adaptation, aes(x = .borders, y = .value, fill=.crop)) +
          geom_bar(stat = "identity") + 
          ylim(ylims) +
          ylab(ylabs) +
          xlab("") + 
          theme_bw() + theme(legend.position = "none") +
          scale_fill_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4])) + 
          guides(colour = guide_legend(override.aes = list(size=5)))
}
  p7 <- ggplot(efdf.adaptation, aes(x = .feature, y = .value, fill=.crop)) +
          geom_bar(stat = "identity") + 
          ylim(ylims) +
          ylab(ylabs) +
          xlab("") + 
          theme_bw() +
          labs(fill="Crop") +
          scale_fill_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4]), name="Crop") +
          guides(colour = guide_legend(override.aes = list(size=5)))

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs.pdf"), height=6, width=6.5)
    multiplot(p1,p2,p3,p8,p4,p5, cols=2)
dev.off()

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs.pdf"), height=6, width=6.5)
    multiplot(p1,p2,p3,p8,p9,p4,p5, cols=2)
dev.off()

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs_adapt.pdf"), height=5, width=3.25)
    p6
dev.off()

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs_legend.pdf"), height=5, width=5)
    p7
dev.off()
```

#### Make PDP 2 way plots
```{r}
same_locations = F
adaptation_types = F
do_iml = T
Nb = 1
run_all_mods = F
train_on_all = T
fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_store_bestML_PDP2way.Rdata"))

PDP2way[[1]][[1]]$results$Crop <- rep("Maize",dim(PDP2way[[1]][[1]]$results)[1])
PDP2way[[1]][[2]]$results$Crop <- rep("Rice",dim(PDP2way[[1]][[2]]$results)[1])
PDP2way[[1]][[3]]$results$Crop <- rep("Wheat",dim(PDP2way[[1]][[3]]$results)[1])
PDP2way[[1]][[4]]$results$Crop <- rep("Soybean",dim(PDP2way[[1]][[4]]$results)[1])

bigpdp2 <- rbind(PDP2way[[1]][[1]]$results, PDP2way[[1]][[2]]$results, PDP2way[[1]][[3]]$results, PDP2way[[1]][[4]]$results)
bigpdp <- bigpdp2 %>% dplyr::select(-c(.type)) %>% pivot_wider(id_cols=c(Crop,Delta_temp), names_from = Adaptation, values_from = .value) %>% mutate(Difference = Yes-No) %>% dplyr::select(-c(No, Yes))

pdf(file=paste0(figoutdir,"PDP2way.pdf"), height=3, width=5)
      ggplot(bigpdp, aes(Delta_temp, Difference, color= Crop)) + geom_line(lwd=2) +
        ylab("Increase in Yield (%) with Adaptation") +
        xlab(expression(Delta~"Temperature"~degree~"C")) + 
        theme_bw() + theme(legend.position = "none") +
        scale_color_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4])) + 
        guides(colour = guide_legend(override.aes = list(size=5)))
dev.off()
```

####Output linear mixed models
```{r}
same_locations = F
adaptation_types = F
do_iml = F
Nb = 1
run_all_mods = T
train_on_all = T
fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_saved_modOutLm.Rdata"))
load(paste0(fileprefix,"_saved_modOutLmInts.Rdata"))
load(paste0(fileprefix,"_saved_modLm.Rdata"))
load(paste0(fileprefix,"_saved_modLmInts.Rdata"))

save.mod.out.lm[[1]][[1]]
save.mod.out.lm[[1]][[2]]
save.mod.out.lm[[1]][[3]]
save.mod.out.lm[[1]][[4]]

save.mod.out.lm.ints[[1]][[1]]
save.mod.out.lm.ints[[1]][[2]]
save.mod.out.lm.ints[[1]][[3]]
save.mod.out.lm.ints[[1]][[4]]

save.mod.lm[[1]]
save.mod.lm.ints[[1]]

summary(save.mod.lm[[1]][[1]])
summary(save.mod.lm[[1]][[2]])
summary(save.mod.lm[[1]][[3]])
summary(save.mod.lm[[1]][[4]])
```


#Shap loop
##Set switches and read data
```{r}
same_locations = F
crop_species = c("Maize","Rice","Wheat","Soybean")
adaptation_types = F
run_all_mods = F
train_on_all = T
rcp45 = T
do_shap = F
do_iml=F
Nb=1

######Load models and data#####
outputdir=paste0(rootdir,"/Output/Rdata/")
figoutdir=paste0(rootdir,"/Output/Figures/")

fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_saved_bestMLs.Rdata")) #save.mod.rf[[1]][[k]]
load(paste0(fileprefix,"_saved_trainingData.Rdata"))

mdatadir <- paste0(rootdir,"/Data/")

TPavedata <- paste0(rootdir, "/MIRCA2000_W5ES_CountryData_fromToshi/annual_area_harvested_ha_30mn_with_climate.xlsx")
Pave <- read.xlsx(TPavedata, sheet="Pave")
Tave <- read.xlsx(TPavedata, sheet="Tave")
CropArea <- read.xlsx(TPavedata, sheet="annual_area_harvested_ha&T&Pr")
```

###Need to load delta precip and delta temp
```{r}
load(paste0(rootdir,"/RFmodels_predicting_tempAndPrecipChange.Rdata"))
#mod.delta.temp.g, mod.delta.temp.l, mod.pr.change, mod.pr.change2

Delta_temp_level = 2
TemplData <- data.frame(Delta_temp_g=rep(Delta_temp_level, nrow(Tave)), 
                        Latitude=Tave$y, 
                        Longitude=Tave$x)
  
  #Prediction RF
  Pred_templ<- predict(mod.delta.temp.l, newdata=TemplData)

TemplData$Delta_temp_l <- Pred_templ

#Merge so that x and y are in the same order for TAB_p and Pave
#Remove duplicate columns id, x, and y
Pave_clean <- Pave[,-c(3:5)]
Pave_merged <- left_join(Tave, Pave_clean)

PrChangeData <- data.frame(Delta_temp_g=rep(Delta_temp_level, nrow(Tave)), 
                           Delta_temp_l=TemplData$Delta_temp_l,
                           Latitude=Tave$y, 
                           Longitude=Tave$x,
                           RCP=as.factor(rep("RCP4.5", nrow(Tave))),
                           Futureyear=rep(2060, nrow(Tave)),
                           Pr_avg=Pave_merged$Prm #mean annual precip in mm
                           )

 #Prediction RF
  levels(PrChangeData$RCP) <- c("RCP4.5", "A1b","A1B", "A1b, B1, and A2", "A1F1, B1" ,                 "A1FI" , "A1FI, A2,B1,B2" , "A1FI, B1" ,"A1FI+B1" , "A2" ,"A2b" , "B1",  "B2" , "C", "Others" , "RCP2.6", "RCP6.0", "RCP8.5" ,"SRES(mean of 6 scenarios)")
  Pred_prchange<- predict(mod.pr.change2, newdata=PrChangeData)
  
PrChangeData$PrChange <- Pred_prchange

head(PrChangeData)

#Add Tave to PrChangeData
PrChangeData$Tave <- Tave$Tave
InputData <- na.omit(PrChangeData)
```

###Optional maps of precip and temp change 
```{r}
#make gridded maps
  wrld <- map_data("world")
  xbreaks <- seq(-180,180,60)
  xlabels <- makelabelsEW(xbreaks)
  ybreaks <- seq(-90,90,30)
  ylabels <- makelabelsNS(ybreaks)
  gg1 <- ggplot() +
    geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = "lightgray", color = "lightgray") +
    coord_fixed(1.3) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15),
          legend.text=element_text(size=12), axis.text = element_text(size=12))+
    scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
    scale_y_continuous("", breaks = ybreaks, labels = ylabels, limits=c(-60,90))


    pdf(file=paste0(figoutdir,"PrAvg.pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=PrChangeData, aes(x=Longitude, y=Latitude, fill=Pr_avg), alpha=0.8) + scale_fill_gradient2(midpoint=0, low=lowcolor, mid=midcolor, high=highcolor, name="Avg Precip (mm)") +
      ggtitle("Average Precipitation (mm)") +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"PrChange.pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=PrChangeData, aes(x=Longitude, y=Latitude, fill=PrChange), alpha=0.8) + scale_fill_gradient2(midpoint=0, low=lowcolor, mid=midcolor, high=highcolor, name="Delta Precip (mm)", limits=c(-300,400)) + 
      ggtitle("Change in Precipitation (mm)") +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"TAvg.pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=PrChangeData, aes(x=Longitude, y=Latitude, fill=Tave), alpha=0.8) + scale_fill_gradient2(midpoint=0, low=lowcolor, mid=midcolor, high=highcolor, name="Avg Temp (degC)") + 
      ggtitle("Average Temperature (degC)") +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
```

##Update Shap loop
```{r}
rcp45 = T
do_shap = T

  Delta_temp_levels = c(2,2)
  CO2.ppm_levels = c(116,116)
  Adaptation_levels = c("No","Yes")

start <- Sys.time()
numbers <- c(1,2,3,4)
for(k in numbers){ 
  crop_species_map = crop_species[[k]]
  #Filter by cropping areas
  Area_crop<-read.csv(paste0(mdatadir,crop_species_map,"_country.csv"), header=T)
  names(Area_crop)<-c("ID","Longitude","Latitude","IRC","RFC","Total","Country","Flag")
  TAB_in<-right_join(Area_crop,InputData, by=c("Latitude","Longitude"))
  TAB_in<-TAB_in[TAB_in$Total*TAB_in$Flag>0,]
  
    mod.rf.map <- save.mod.rf[[1]][[k]]
    Training_map <- save.Training.rf[[1]][[k]]

    for(levelindex in 1:length(Delta_temp_levels)){
      do_shaps(Delta_temp_levels[levelindex], CO2.ppm_levels[levelindex], Adaptation_levels[levelindex], mod.rf.map, do_shap, TAB_in, levelindex, k)
    }
   
}
end <- Sys.time()
save.time <- end - start
```

###Plot Shapley Maps
```{r}
shap_df <- NULL
gg1 <- makebaseplot()
cbPalette <- c("#000000","#E69F00", "#009E73", "#0072B2", "#CC79A7", "#D55E00", "#56B4E9", "#F0E442")

for(AdaptLevel in 1:2){
for(k in 1:length(crop_species)){

load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_co2ppm_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_TempAvg_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_Delta_temp_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_Adaptation_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_Precip_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_Delta_precip_4.5.RData"))

#Filter by cropping areas
  Area_crop<-read.csv(paste0(mdatadir,crop_species[[k]],"_country.csv"), header=T)
  names(Area_crop)<-c("ID","Longitude","Latitude","IRC","RFC","Total","Country","Flag")
  TAB_in<-right_join(Area_crop,InputData, by=c("Latitude","Longitude"))
  TAB_in<-TAB_in[TAB_in$Total*TAB_in$Flag>0,]

      map_df <- map_of_scenario(AdaptLevel)

      shap_df[[k]] <- map_df %>% tidyr::pivot_wider(names_from="Variable",values_from="Effect")
      
map_df$Variable <- factor(map_df$Variable, 
                        levels=c("Adaptation", "co2ppm", "Delta_precip", "Delta_temp", "Precip", "TempAvg"),
                        labels=c("Adaptation", expression(Delta~"CO"[2]~"ppm above 390 ppm"), expression(Delta~"Precipitation (mm)"), expression(Delta~"Temperature"~degree~"C"), expression("Average Precipitation"~"(mm)"), expression("Average Temperature"~degree~"C")))

    pdf(file=paste0(figoutdir,"shap_rcp45_",crop_species[k],"_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=map_df, aes(x=x, y=y, fill=Effect), alpha=0.8) + 
      scale_fill_gradientn(colours=c("#533600", "#966D12", "#CBA46C",  "#F6F6F6", "#2EBEB0", "#00897B", "#004B40"), limits=c(-35,35),  breaks=c(-35,-15,0,15,35), name="Yield Change (%)") + 
      ggtitle(paste0(crop_species[k])) +
      facet_wrap(Variable ~ ., nrow=2, labeller = label_parsed) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
    
#Max abs Shap value plots
absshaps <- cbind(abs(shap_TempAvg), abs(shap_Delta_temp), abs(shap_co2ppm), abs(shap_Adaptation), abs(shap_Precip), abs(shap_Delta_precip))
shaps <- cbind(shap_TempAvg, shap_Delta_temp, shap_co2ppm, shap_Adaptation, shap_Precip, shap_Delta_precip)
colnames(absshaps) <- c("Temperature","DeltaTemp","CO2ppm","Adaptation","Precipitation","DeltaPrecip")
colnames(shaps) <- c("Temperature","DeltaTemp","CO2ppm","Adaptation","Precipitation","DeltaPrecip")

shap_df[[k]]$maxabsshaps <- colnames(absshaps)[apply(absshaps,1,which.max)]
shap_df[[k]]$maxshaps <- colnames(shaps)[apply(shaps,1,which.max)]
shap_df[[k]]$minshaps <- colnames(shaps)[apply(shaps,1,which.min)]
shap_df[[k]]$Crop <- rep(crop_species[k], dim(shap_df[[k]])[1])

pdf(file=paste0(figoutdir,"shaphist_",crop_species[k],"_Adapt",AdaptLevel,".pdf"), height=5, width=5)
par(mfrow=c(2,2))
hist(shap_TempAvg, main=expression("Average Temperature"~degree~"C"), xlab="Shapley Value")
hist(shap_Delta_temp, main=expression(Delta~"Temperature"~degree~"C"), xlab="Shapley Value")
hist(shap_co2ppm, main=expression(Delta~"CO"[2]~"ppm above 390 ppm"), xlab="Shapley Value")
hist(shap_Adaptation, main=expression("Adaptation"), xlab="Shapley Value")
hist(shap_Precip, main=expression("Average Precipitation (mm)"), xlab="Shapley Value")
hist(shap_Delta_precip, main=expression(Delta~"Precipitation (mm)"), xlab="Shapley Value")
dev.off()

}

shap_df_map <- do.call("rbind", shap_df)
var.labs <- c(expression("Adaptation"), expression(Delta~"CO"[2]~"ppm"), expression(Delta~"Precipitation (mm)"), expression(Delta~"Temperature"~degree~"C"), expression("Average Precipitation (mm)"), expression("Average Temperature"~degree~"C"))

    pdf(file=paste0(figoutdir,"shapmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=shap_df_map, aes(x=x, y=y, fill=maxshaps), alpha=0.8) + scale_fill_manual(values = c(cbPalette[5], cbPalette[3], cbPalette[4], cbPalette[6], cbPalette[7], cbPalette[8]), labels=var.labs) +  
      facet_wrap(Crop ~ ., nrow=2) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_blank()) #change font size of legend title
    )
    dev.off()
    
      pdf(file=paste0(figoutdir,"barplot_shapmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(ggplot(shap_df_map, aes(x=maxshaps)) + geom_bar() + scale_fill_manual(values = c(cbPalette[5], cbPalette[3], cbPalette[4], cbPalette[6], cbPalette[7], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), 
      axis.text = element_text(size=12),
      axis.text.x=element_text(size=8, angle = 45, vjust = 0.5),
      axis.text.y=element_text(size=8), 
      plot.title=element_text(size=8), 
      legend.text=element_text(size=8), 
      legend.title=element_blank()) +
      ylab("Count") +
      xlab("") +
  scale_x_discrete("", labels = var.labs)
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"shapmin_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=shap_df_map, aes(x=x, y=y, fill=minshaps), alpha=0.8) + scale_fill_manual(values = c(cbPalette[5], cbPalette[3], cbPalette[4], cbPalette[6], cbPalette[7], cbPalette[8]), labels=var.labs) +
      facet_wrap(Crop ~ ., nrow=2) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_blank()) #change font size of legend title
    )
    dev.off()
    
      pdf(file=paste0(figoutdir,"barplot_shapmin_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(ggplot(shap_df_map, aes(x=minshaps)) + geom_bar() + scale_fill_manual(values = c(cbPalette[5], cbPalette[3], cbPalette[4], cbPalette[6], cbPalette[7], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), 
      axis.text = element_text(size=12),
      axis.text.x=element_text(size=8, angle = 45, vjust = 0.5),
      axis.text.y=element_text(size=8), 
      plot.title=element_text(size=8), 
      legend.text=element_text(size=8), 
      legend.title=element_blank()) +
      ylab("Count") +
      xlab("") +
  scale_x_discrete("", labels = var.labs)
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"shapabsmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=shap_df_map, aes(x=x, y=y, fill=maxabsshaps), alpha=0.8) + scale_fill_manual(values = c(cbPalette[5], cbPalette[3], cbPalette[4], cbPalette[6], cbPalette[7], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_blank()) #change font size of legend title
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"barplot_shapabsmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(ggplot(shap_df_map, aes(x=maxabsshaps)) + geom_bar() + scale_fill_manual(values = c(cbPalette[5], cbPalette[3], cbPalette[4], cbPalette[6], cbPalette[7], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), 
      axis.text = element_text(size=12),
      axis.text.x=element_text(size=8, angle = 45, vjust = 0.5),
      axis.text.y=element_text(size=8), 
      plot.title=element_text(size=8), 
      legend.text=element_text(size=8), 
      legend.title=element_blank()) +
      ylab("Count") +
      xlab("") +
  scale_x_discrete("", labels = var.labs)
    )
    dev.off()
    
save(shap_df_map, file=paste0(outputdir,AdaptLevel,"_ShapdfMap.RData"))
}
```

#Fig2 Yield Change Summary Map
```{r}
map_df_adapt <- NULL
map_df_noadapt <- NULL
map_df_all <- NULL

for (k in 1:4){
  load(file=paste0(outputdir,crop_species[k], "_Nb",1,"_Adapt",1,"_predicted_effect_4.5.RData"))
  
  map_df_noadapt[[k]] <- NewData
  map_df_noadapt[[k]]$Crop <- crop_species[k]
  
  load(file=paste0(outputdir,crop_species[k], "_Nb",1,"_Adapt",2,"_predicted_effect_4.5.RData"))
  
  map_df_adapt[[k]] <- NewData
  map_df_adapt[[k]]$Crop <- crop_species[k]
}

map_df_noadapt[[2]]$Adaptation <- rep("No", dim(map_df_noadapt[[2]])[1])
map_df_noadapt[[4]]$Adaptation <- rep("No", dim(map_df_noadapt[[4]])[1])
map_df_adapt[[2]]$Adaptation <- rep("Yes", dim(map_df_adapt[[2]])[1])
map_df_adapt[[4]]$Adaptation <- rep("Yes", dim(map_df_adapt[[4]])[1])

map_df_all <- map_df_adapt[[1]] %>% full_join(map_df_adapt[[2]]) %>% full_join(map_df_adapt[[3]]) %>% full_join(map_df_adapt[[4]]) %>% full_join(map_df_noadapt[[1]]) %>% full_join(map_df_noadapt[[2]]) %>% full_join(map_df_noadapt[[3]]) %>% full_join(map_df_noadapt[[4]]) 

# New facet label names for Adaptation variable
adpt.labs <- c("No Adaptation", "Adaptation")
names(adpt.labs) <- c("No", "Yes")

adpt.labs.change <- c("No Adaptation", "Adaptation","Difference")
names(adpt.labs.change) <- c("No", "Yes", "Change")

map_df_all$Crop <- as.factor(map_df_all$Crop)

map_change <- map_df_all %>% tidyr::pivot_wider(names_from=Adaptation, values_from=Effect) %>% mutate(Change = Yes-No) %>% tidyr::pivot_longer(cols=c(Yes, No, Change), names_to = "Adaptation", values_to = "Effect") 

map_change$Adaptation_f <- factor(map_change$Adaptation, levels=c("No","Yes","Change"))

rcpheight <- 7
rcpwidth <- 9
rcpsize <- 8
###Plot####
    pdf(file=paste0(figoutdir,"allcrops_rcp45",rcp45,".pdf"), height=rcpheight, width=rcpwidth)
    print(gg1 + geom_tile(data=map_df_all, aes(x=Longitude, y=Latitude, fill=Effect), alpha=0.8) + 
      scale_fill_gradientn(colours=c("#533600", "#966D12", "#CBA46C",  "#F6F6F6", "#2EBEB0", "#00897B", "#004B40"), limits=c(-65,65),  breaks=c(-65,-25,0,25,65), name="Yield Change (%)") +
      ggtitle(paste0("RCP 4.5")) +
      facet_grid(Adaptation ~ Crop, labeller = labeller(Adaptation = adpt.labs)) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=rcpsize, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=rcpsize), #change font size of axis titles
        plot.title=element_text(size=rcpsize), #change font size of plot title
        legend.text=element_text(size=rcpsize), #change font size of legend text
        legend.title=element_text(size=rcpsize)) #change font size of legend title
    )
    dev.off()
    
    pdf(file=paste0(figoutdir,"allcrops_rcp45",rcp45,"_change.pdf"), height=rcpheight, width=rcpwidth)
    print(gg1 + geom_tile(data=map_change, aes(x=Longitude, y=Latitude, fill=Effect), alpha=0.8) + 
        scale_fill_gradientn(colours=c("#533600", "#966D12", "#CBA46C",  "#F6F6F6", "#2EBEB0", "#00897B", "#004B40"), limits=c(-65,65),  breaks=c(-65,-25,0,25,65), name="Yield Change (%)") + 
      ggtitle(paste0("RCP 4.5")) +
      facet_grid(Adaptation_f ~ Crop, labeller = labeller(Adaptation_f = adpt.labs.change)) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=rcpsize, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=rcpsize), #change font size of axis titles
        plot.title=element_text(size=rcpsize), #change font size of plot title
        legend.text=element_text(size=rcpsize), #change font size of legend text
        legend.title=element_text(size=rcpsize)) #change font size of legend title
    )
    dev.off()
    
      pdf(file=paste0(figoutdir,"allcrops_rcp45",rcp45,"_boxplot.pdf"), height=3, width=5)
          print(ggplot(data=map_df_all, aes(x=Crop, y=Effect, fill=Adaptation), alpha=0.8) + 
                  geom_boxplot(outlier.shape = NA) + stat_boxplot(geom ='errorbar', coef=NULL) +
                    scale_fill_manual(values=c("darkgray","lightgray")) +
                   ggtitle(paste0("RCP 4.5")) + theme_classic() +
                  geom_hline(yintercept=0, linetype='dotted') +
              ylab("Yield Change (%)") #change font size of legend title
          )
      dev.off()

#Statistically significant effect of adaptation
stat_test <- lm(map_df_all$Effect ~ map_df_all$Adaptation + map_df_all$Crop)
summary(stat_test)
summary.aov(stat_test)
plot(stat_test)
```

# Load shapmaps and summaries
```{r}
load(paste0(outputdir,"1_ShapdfMap.RData"))
adpt1shap <- shap_df_map
load(paste0(outputdir,"2_ShapdfMap.RData"))
adpt2shap <- shap_df_map

#Summaries
table(adpt1shap$maxabsshaps, by=adpt1shap$Crop)
table(adpt1shap$maxshaps, by=adpt1shap$Crop)
table(adpt1shap$minshaps, by=adpt1shap$Crop)

table(adpt2shap$maxabsshaps, by=adpt2shap$Crop)
table(adpt2shap$maxshaps, by=adpt2shap$Crop)
table(adpt2shap$minshaps, by=adpt2shap$Crop)

summary(adpt1shap[adpt1shap$Crop=="Maize",]$Adaptation)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$Adaptation)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$TempAvg)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$TempAvg)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$Precip)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$Precip)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$Delta_temp)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$Delta_temp)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$co2ppm)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$co2ppm)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$Delta_precip)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$Delta_precip)

summary(adpt1shap[adpt1shap$Crop=="Rice",]$Adaptation)
summary(adpt2shap[adpt2shap$Crop=="Rice",]$Adaptation)
summary(adpt1shap[adpt1shap$Crop=="Rice",]$TempAvg)
summary(adpt2shap[adpt2shap$Crop=="Rice",]$TempAvg)

summary(adpt1shap[adpt1shap$Crop=="Soybean",]$Adaptation)
summary(adpt2shap[adpt2shap$Crop=="Soybean",]$Adaptation)
summary(adpt1shap[adpt1shap$Crop=="Soybean",]$TempAvg)
summary(adpt2shap[adpt2shap$Crop=="Soybean",]$TempAvg)
```

# Shap Value Boxplots
```{r}
shap_df_all1 <- adpt1shap %>% dplyr::select(Delta_precip, Precip, TempAvg, Delta_temp, co2ppm, Adaptation, Crop) %>% tidyr::pivot_longer(cols=c(Delta_precip, Precip, TempAvg, Delta_temp, co2ppm, Adaptation), names_to = "Variable", values_to = "Effect")
shap_df_all2 <- adpt2shap %>% dplyr::select(Delta_precip, Precip, TempAvg, Delta_temp, co2ppm, Adaptation, Crop) %>% tidyr::pivot_longer(cols=c(Delta_precip, Precip, TempAvg, Delta_temp, co2ppm, Adaptation), names_to = "Variable", values_to = "Effect")
      
    pdf(file=paste0(figoutdir,"shap_box_Adapt1.pdf"), height=4, width=6)
              print(ggplot(data=shap_df_all1, aes(x=Variable, y=Effect)) +
                  geom_boxplot(outlier.shape = NA) + stat_boxplot(geom ='errorbar', coef=NULL) +
                  theme_classic() +
                  geom_hline(yintercept=0, linetype='dotted') +
                  ylab("Shapley Value") +
                  facet_wrap(Crop ~ ., nrow=2) + xlab("") +
                   scale_x_discrete(breaks=c("Adaptation", "co2ppm", "Delta_precip", "Delta_temp", "Precip", "TempAvg"), labels=c("No Adaptation", expression(Delta~"CO"[2]), expression(Delta~"Precip"), expression(Delta~"Temp"), expression("Avg Precip"), expression("Avg Temp"))) +
                  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)))
    dev.off()
    
    pdf(file=paste0(figoutdir,"shap_box_Adapt2.pdf"), height=4, width=6)
              print(ggplot(data=shap_df_all2, aes(x=Variable, y=Effect)) +
                  geom_boxplot(outlier.shape = NA) + stat_boxplot(geom ='errorbar', coef=NULL) +
                  theme_classic() +
                  geom_hline(yintercept=0, linetype='dotted') +
                  ylab("Shapley Value") +
                  facet_wrap(Crop ~ ., nrow=2) + xlab("") +
                    scale_x_discrete(breaks=c("Adaptation", "co2ppm", "Delta_precip","Delta_temp", "Precip", "TempAvg"), labels=c("Adaptation", expression(Delta~"CO"[2]), expression(Delta~"Precip"), expression(Delta~"Temp"), expression("Avg Precip"), expression("Avg Temp"))) +
                  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)))
    dev.off()
```


#Summaries
```{r}
######Summary stats#####
summary_stats <- map_df_all %>%
  group_by(Delta_temp, CO2.ppm, Adaptation, Crop) %>%
  dplyr::summarise(Mean = mean(Effect), 
                   Median = median(Effect), 
                   SE = sd(Effect)/sqrt(n()),
                   min = min(Effect),
                   max = max(Effect),
                   n = n())
summary_stats
```


