---
title: "Yield_analysis"
output: html_document
---
```
Initial ML workflow drafted by David Makowski INRAE (18 Feb 2021)
Additional ML analyses, Shapley explanations, and Figures developed by Rose Abramoff (22 Feb 2021)

This analysis has two main sections:
1) testing the performance of different random forest models and generating plots
2) calculating Shapley additive explanations and generating plots

The analysis sources functions:
multiplot.R (for plotting)
readin.R (for reading in data)
plotfuncs.R (for plotting)
yieldloop.R (for testing ML models)
shaploop.R (for calculating Shapley values)
```
#Load libraries
```{r load-libraries}
rm(list=ls()) 

library(openxlsx)
library(dplyr)
library(ggplot2)
library(spaMM)
library(ranger)
library(gbm)
library(iml)
library(lme4)
library(randomForest)
library(mapdata)
library(raster)
library(DALEX)
```

#Switches
```{r switches}
do_maps = T
```

#Data import
```{r data-import}
rootdir <- "/Users/r4y/Dropbox (ORNL)/Collaborations/Makowski_yield"
outputdir= paste0(rootdir,"/Output/Rdata/")
figoutdir= paste0(rootdir,"/Output/Figures/")
crop_species = c("Maize","Rice","Wheat","Soybean")
source('multiplot.R')
source('readin.R')
source('plotfuncs.R')
source('yieldloop.R')
source('shaploop.R')
FTAB <- readin()
```

#Do Maps
```{r, do-maps}
if(do_maps){
pdf(file=paste0(figoutdir,"yield_meta.pdf"), height=8, width=10)
print(map_of_df(FTAB))
dev.off()

pdf(file=paste0(figoutdir,"yield_hist.pdf"), height=6, width=6)
print(hist_of_df(FTAB))
dev.off()
}
```

#Describe data
```{r}
FTAB %>% 
  group_by(Crop) %>% 
  summarize(locations = length(unique(Loca)),
            samples = length(Effect), #complete.cases does not change value
            meanEffect = mean(Effect),
            medianEffect = median(Effect),
            minEffect = min(Effect),
            maxEffect = max(Effect),
            sd = sd(Effect),
            se = sd(Effect)/sqrt(length(Effect)))

# View(FTAB %>%
#     group_by(Crop,Adaptation.type) %>%
#   summarize(n()))

shorter <- FTAB %>%
  dplyr::select(Effect,Crop,Delta_temp,CO2.ppm,TempAvg,Precip,Delta_precip,Latitude,Longitude,Fertiliser,Irrigation,Cultivar,Soil.organic.matter.management,Planting.time,Tillage,Others,Loca) %>%
  tidyr::drop_na()

length(unique(shorter$Loca))
```

#Yield Loop
```{r, eval=F}
#For training
start <- Sys.time()
loop_func(same_locations = T, adaptation_types = F, do_iml = F, Nb = 10, run_all_mods = T, train_on_all = F)
end <- Sys.time()
end-start
#Approx 45 mins

start <- Sys.time()
loop_func(same_locations = F, adaptation_types = F, do_iml = F, Nb = 10, run_all_mods = T, train_on_all = F)
end <- Sys.time()
end-start
#Approx 30 mins

#For training
start <- Sys.time()
loop_func(same_locations = T, adaptation_types = F, do_iml = T, Nb = 10, run_all_mods = F, train_on_all = F)
end <- Sys.time()
end-start
#Approx 2.6 hr

start <- Sys.time()
loop_func(same_locations = F, adaptation_types = F, do_iml = T, Nb = 10, run_all_mods = F, train_on_all = F)
end <- Sys.time()
end-start

same_locations = F
adaptation_types = T
do_iml = T
Nb = 1
run_all_mods = F
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start

same_locations = F
adaptation_types = T
do_iml = F
Nb = 1
run_all_mods = F
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start

same_locations = F
adaptation_types = F
do_iml = F
Nb = 1
run_all_mods = F
train_on_all = T
start <- Sys.time()
loop_func(same_locations, adaptation_types, do_iml, Nb, run_all_mods, train_on_all)
end <- Sys.time()
end-start
```

##Make Plots
###Make obs v pred pred plots
```{r}
same_locations = T #changed - run this combo
adaptation_types = F
do_iml = F
Nb = 10
run_all_mods = T
train_on_all = F
fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_store_bestML_predictions.Rdata"))

pdf(file=paste0(figoutdir,"FigureS3_samelocs",same_locations,"_adaptypes",adaptation_types,".pdf"), height=4, width=6)
obs_v_pred(ObsPred)
dev.off()

resultdf <- read.csv(paste0(fileprefix,"_results.csv"))
resultdf

  AllOP <- bind_rows(ObsPred, .id = "column_label")
  FormOP <- AllOP %>%
    tidyr::pivot_longer(cols=c(Predicted, Observed), names_to="Type") %>%
    group_by(Crop, Site, Type) %>%
    
    dplyr::summarise(mean = mean(value),
                     se = sd(value)/sqrt(n()),
                     n = n()) %>%
    tidyr::pivot_wider(names_from = Type, names_sep = ".", values_from = c(mean, se, n)) %>% mutate(Residual=mean.Observed - mean.Predicted)
  
pdf(file=paste0(figoutdir,"Residuals_samelocs",same_locations,"_adaptypes",adaptation_types,".pdf"), height=4, width=6)
  plot(FormOP$mean.Predicted, FormOP$Residual, ylab="Residuals", xlab="Observed Yield Change")
dev.off()
```

###Make PDP plots
####Load the data
```{r}
same_locations = F
adaptation_types = T
do_iml = T
Nb = 1
run_all_mods = F
train_on_all = T
fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_store_bestML_PDPs.Rdata"))
#PDPeffs[[Nb 1-10]][[Crop 1-4]]
```

####Make the plots
```{r}
cbPalette <- c("#000000","#E69F00", "#009E73", "#0072B2", "#CC79A7", "#D55E00", "#56B4E9", "#F0E442")

bindeffs <- function(PDPeffs, Crop.Num) {
  effs <- NULL
for(i in 1:1){
effs[[i]] <- do.call("rbind", PDPeffs[[i]][[Crop.Num]]$results)
effs[[i]]$.crop <- rep(crop_species[Crop.Num], dim(effs[[i]])[1])
}
neweffs <- do.call("rbind", effs)
return(neweffs)
}

maize.effs <- bindeffs(PDPeffs, 1)
rice.effs <- bindeffs(PDPeffs, 2)
wheat.effs <- bindeffs(PDPeffs, 3)
soybean.effs <- bindeffs(PDPeffs, 4)

     efdf <- maize.effs %>% 
       dplyr::full_join(rice.effs) %>%
       dplyr::full_join(wheat.effs) %>%
       dplyr::full_join(soybean.effs) %>%
        group_by(.borders, .feature, .crop) 
     efdf$.crop <- as.factor(efdf$.crop)
     
     #levels(efdf$.crop) <- c("Maize","Rice","Wheat","Soybean")
     
if(adaptation_types){
    efdf.adaptation <- filter(efdf, .feature != "TempAvg" && .feature != "CO2.ppm" && .feature != "Delta_temp" && .feature != "Latitude" && .feature != "Longitude" && .feature != "Precip" && .feature != "Delta_precip")
    efdf.others <- filter(efdf, .feature == "TempAvg" | .feature == "CO2.ppm" | .feature == "Delta_temp" | .feature == "Latitude" | .feature == "Longitude" | .feature == "Precip" | .feature == "Delta_precip")
    
    # efdf.adaptation <- filter(efdf, .feature != "TempAvg" && .feature != "CO2.ppm" && .feature != "Delta_temp" && .feature != "Latitude" && .feature != "Longitude" && .feature != "Precip")
    # efdf.others <- filter(efdf, .feature == "TempAvg" | .feature == "CO2.ppm" | .feature == "Delta_temp" | .feature == "Latitude" | .feature == "Longitude" | .feature == "Precip")
    
    efdf.adaptation$.feature <- as.factor(efdf.adaptation$.feature)
    levels(efdf.adaptation$.feature) <- c("Cultivar", "Fertilizer", "Irrigation", "Others", "Planting Time")
    efdf.adaptation <- efdf.adaptation %>%
      tidyr::pivot_wider(names_from=.borders, values_from=.value) %>%
      mutate(.value = Yes-No)
}else{
    efdf.adaptation <- efdf[efdf$.feature=="Adaptation",]
    efdf.others <- efdf[efdf$.feature!="Adaptation",]
}
     
     efdf.others$.borders <- as.numeric(efdf.others$.borders)
     efdf.TempAvg <- efdf.others[efdf.others$.feature=="TempAvg",]
     efdf.Delta_temp <- efdf.others[efdf.others$.feature=="Delta_temp",]
     efdf.CO2.ppm <- efdf.others[efdf.others$.feature=="CO2.ppm",]
     efdf.Latitude <- efdf.others[efdf.others$.feature=="Latitude",]
     efdf.Longitude <- efdf.others[efdf.others$.feature=="Longitude",]
     efdf.Precip <- efdf.others[efdf.others$.feature=="Precip",]
     
     efdf.Delta_precip <- efdf.others[efdf.others$.feature=="Delta_precip",]
    
      ylabs = "Yield Change (%)"
      ylims = range(efdf$.value)+range(efdf$.value)*0.1
      
      ggplot_pdps <- function(efdf, xlabels) {
      ggplot(efdf, aes(.borders, .value, color= .crop)) +
        geom_point(alpha=0.3) + geom_line() +
        ylim(ylims) +
        ylab(ylabs) +
        xlab(xlabels) + 
        theme_bw() + theme(legend.position = "none") +
        #scale_color_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4]), labels = c("Maize", "Rice",  "Wheat", "Soybean")) + 
        scale_color_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4])) + 
        guides(colour = guide_legend(override.aes = list(size=5)))
      }
      
      ggplot_pdps_latlong <- function(efdf, xlabels) {
      ggplot(efdf, aes(.borders, .value, color= .crop)) +
        geom_point(alpha=0.3) + geom_line() +
        ylim(ylims) +
        ylab(ylabs) +
        xlab(xlabels) + 
        theme_bw() + theme(legend.position = "none") +
        #scale_color_manual(values = c(cbPalette[2], cbPalette[4])) + 
        scale_color_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4])) + 
        guides(colour = guide_legend(override.aes = list(size=5)))
      }
      
p1 <- ggplot_pdps(efdf.TempAvg, expression("Average Temperature"~degree~"C"))
p2 <- ggplot_pdps(efdf.Delta_temp, expression(Delta~" Temperature"~degree~"C"))
p3 <- ggplot_pdps(efdf.CO2.ppm, expression(Delta~"CO"[2]~"ppm above 390 ppm"))
p4 <- ggplot_pdps_latlong(efdf.Latitude, "Latitude")
p5 <- ggplot_pdps_latlong(efdf.Longitude, "Longitude")
p8 <- ggplot_pdps(efdf.Precip, expression("Annual Precipitation (mm)"))

p9 <- ggplot_pdps(efdf.Delta_precip, expression("Annual "~Delta~" Precipitation (mm)"))

if(adaptation_types){
  p6 <- ggplot(efdf.adaptation, aes(x = .feature, y = .value, fill=.crop)) +
          geom_bar(stat = "identity", position="dodge") + 
          facet_grid(.crop~.) +
          #ylim(-16,11) +
          ylab("Yield Improvement with Adaptation (%)") +
          xlab("") + 
          theme_bw() + theme(legend.position = "none", strip.text.x = element_text(size = 5), axis.text.x=element_text(angle = 45, hjust= 1)) +
          scale_fill_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4])) + 
          guides(colour = guide_legend(override.aes = list(size=5)))
}else{
  p6 <- ggplot(efdf.adaptation, aes(x = .borders, y = .value, fill=.crop)) +
          geom_bar(stat = "identity") + 
          ylim(ylims) +
          ylab(ylabs) +
          xlab("") + 
          theme_bw() + theme(legend.position = "none") +
          scale_fill_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4])) + 
          guides(colour = guide_legend(override.aes = list(size=5)))
}
  p7 <- ggplot(efdf.adaptation, aes(x = .feature, y = .value, fill=.crop)) +
          geom_bar(stat = "identity") + 
          ylim(ylims) +
          ylab(ylabs) +
          xlab("") + 
          theme_bw() +
          labs(fill="Crop") +
          scale_fill_manual(values = c(cbPalette[1], cbPalette[2], cbPalette[3], cbPalette[4]), name="Crop") +
          guides(colour = guide_legend(override.aes = list(size=5)))

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs.pdf"), height=6, width=6.5)
    multiplot(p1,p2,p3,p8,p4,p5, cols=2)
dev.off()

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs.pdf"), height=6, width=6.5)
    multiplot(p1,p2,p3,p8,p9,p4,p5, cols=2)
dev.off()

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs_adapt.pdf"), height=6, width=3.25)
    p6
dev.off()

pdf(file=paste0(figoutdir,"adaptation",adaptation_types,"_PDPs_legend.pdf"), height=5, width=5)
    p7
dev.off()
```

#Shap loop
##Set switches and read data
```{r}
same_locations = F
crop_species = c("Maize","Rice","Wheat","Soybean")
adaptation_types = F
run_all_mods = F
train_on_all = T
rcp45 = T
do_shap = F
do_iml=F
Nb=1

######Load models and data#####
outputdir="/Users/r4y/Dropbox (ORNL)/Collaborations/Makowski_yield/Output/Rdata/"
figoutdir="/Users/r4y/Dropbox (ORNL)/Collaborations/Makowski_yield/Output/Figures/"

fileprefix <- paste0(outputdir,"sameLocs",same_locations,"_adaptationTypes",adaptation_types,"_doIML",do_iml,"_Nb",Nb,"runAllMods",run_all_mods,"_trainOnAll",train_on_all)
load(paste0(fileprefix,"_saved_bestMLs.Rdata")) #save.mod.rf[[1]][[k]]
#save.mod.rf[[1]][[1]] #Climate only (TempAvg, Delta_temp, Precip, CO2.ppm, Delta_precip, Adaptation)
#save.mod.rf[[1]][[2-4]] #Climate + longlat (Latitude, Longitude, TempAvg, Delta_temp, Precip, CO2.ppm, Delta_precip, Adaptation)
load(paste0(fileprefix,"_saved_trainingData.Rdata"))

mdatadir <- "/Users/r4y/Dropbox (ORNL)/Collaborations/Makowski_yield/Data/"
# TAB_p<-read.csv(paste0(mdatadir,"tave_201120.csv"), header=T)
# TAB_p<-na.omit(TAB_p)

TPavedata <- paste0(rootdir, "/MIRCA2000_W5ES_CountryData_fromToshi/annual_area_harvested_ha_30mn_with_climate.xlsx")
Pave <- read.xlsx(TPavedata, sheet="Pave")
Tave <- read.xlsx(TPavedata, sheet="Tave")
CropArea <- read.xlsx(TPavedata, sheet="annual_area_harvested_ha&T&Pr")
```

###Need to load delta precip and delta temp
```{r}
load(paste0(rootdir,"/RFmodels_predicting_tempAndPrecipChange.Rdata"))
#mod.delta.temp.g, mod.delta.temp.l, mod.pr.change, mod.pr.change2

Delta_temp_level = 2
TemplData <- data.frame(Delta_temp_g=rep(Delta_temp_level, nrow(Tave)), 
                        Latitude=Tave$y, 
                        Longitude=Tave$x)
  
  #Prediction RF
  Pred_templ<- predict(mod.delta.temp.l, newdata=TemplData)

#but then x and y's are not in order
TemplData$Delta_temp_l <- Pred_templ

#Merge so that x and y are in the same order for TAB_p and Pave
#Remove duplicate columns id, x, and y
Pave_clean <- Pave[,-c(3:5)]
Pave_merged <- left_join(Tave, Pave_clean)

PrChangeData <- data.frame(Delta_temp_g=rep(Delta_temp_level, nrow(Tave)), 
                           Delta_temp_l=TemplData$Delta_temp_l,
                           Latitude=Tave$y, 
                           Longitude=Tave$x,
                           RCP=as.factor(rep("RCP4.5", nrow(Tave))),
                           Futureyear=rep(2060, nrow(Tave)),
                           Pr_avg=Pave_merged$Prm #mean annual precip in mm #missing some values
                           )

 #Prediction RF
  levels(PrChangeData$RCP) <- c("RCP4.5", "A1b","A1B", "A1b, B1, and A2", "A1F1, B1" ,                 "A1FI" , "A1FI, A2,B1,B2" , "A1FI, B1" ,"A1FI+B1" , "A2" ,"A2b" , "B1",  "B2" , "C", "Others" , "RCP2.6", "RCP6.0", "RCP8.5" ,"SRES(mean of 6 scenarios)")
  Pred_prchange<- predict(mod.pr.change2, newdata=PrChangeData)
  
PrChangeData$PrChange <- Pred_prchange

head(PrChangeData)

#Now just add Tave to PrChangeData then can move on to shaps
#Tavealt is the same as TAB_p
PrChangeData$Tave <- Tave$Tave
InputData <- na.omit(PrChangeData)

#Filter by cropping areas #Both (Area_crop and CropArea) files are the same, except that 4 separate files that David sent me also has a "Flag" variable which decreases the total number of samples by about 12k
  #Area_crop<-read.csv(paste0(mdatadir,crop_species[[1]],"_country.csv"), header=T)
  #TAB_p_f<-Area_crop[Area_crop$Total*Area_crop$Flag>0,]
```

###Optional maps of precip and temp change 
```{r}
# ## Nice map of precip and temp change
# prdf <- read.csv(paste0(rootdir,"/Predicted_Pr_change_RF.csv"))
# tgdf <- read.csv(paste0(rootdir,"/Predicted_global_temp_change_RF.csv"))
# tldf <- read.csv(paste0(rootdir,"/Predicted_local_temp_change_RF.csv"))
# 
#   wrld <- map_data("world")
#   xbreaks <- seq(-180,180,60)
#   xlabels <- makelabelsEW(xbreaks)
#   ybreaks <- seq(-90,90,30)
#   ylabels <- makelabelsNS(ybreaks)
#   gg1 <- ggplot() +
#     geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = "lightgray", color = "lightgray") +
#     coord_fixed(1.3) +
#     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#           panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15),
#           legend.text=element_text(size=12), axis.text = element_text(size=12))+
#     scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
#     scale_y_continuous("", breaks = ybreaks, labels = ylabels, limits=c(-60,90))
# 
# pdf(file=paste0(figoutdir,"precip_change_points.pdf"), height=6, width=9)
#   gg1 + geom_point(data=prdf, aes(x=Longitude, y=Latitude, color=Pr_avg), pch=16, size=1) +
#     labs(color="Pr_change")
# dev.off()
# 
# pdf(file=paste0(figoutdir,"global_temp_change_points.pdf"), height=6, width=9)
#   gg1 + geom_point(data=tgdf, aes(x=Longitude, y=Latitude, color=Delta_temp_g_pred), pch=16, size=1) +
#     labs(color="Tg_change")
# dev.off()
# 
# pdf(file=paste0(figoutdir,"local_temp_change_points.pdf"), height=6, width=9)
#   gg1 + geom_point(data=tldf, aes(x=Longitude, y=Latitude, color=Delta_temp_l_pred), pch=16, size=1) +
#     labs(color="Tl_change")
# dev.off()

#make gridded maps
  wrld <- map_data("world")
  xbreaks <- seq(-180,180,60)
  xlabels <- makelabelsEW(xbreaks)
  ybreaks <- seq(-90,90,30)
  ylabels <- makelabelsNS(ybreaks)
  gg1 <- ggplot() +
    geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = "lightgray", color = "lightgray") +
    coord_fixed(1.3) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15),
          legend.text=element_text(size=12), axis.text = element_text(size=12))+
    scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
    scale_y_continuous("", breaks = ybreaks, labels = ylabels, limits=c(-60,90))


    pdf(file=paste0(figoutdir,"PrAvg.pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=PrChangeData, aes(x=Longitude, y=Latitude, fill=Pr_avg), alpha=0.8) + scale_fill_gradient2(midpoint=0, low="#440154FF", mid="#21908CFF", high="#FDE725FF", name="Avg Precip (mm)") + #, limits=c(-100,100)) + 
      ggtitle("Average Precipitation (mm)") +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"PrChange.pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=PrChangeData, aes(x=Longitude, y=Latitude, fill=PrChange), alpha=0.8) + scale_fill_gradient2(midpoint=0, low="#440154FF", mid="#21908CFF", high="#FDE725FF", name="Delta Precip (mm)", limits=c(-300,400)) + 
      ggtitle("Change in Precipitation (mm)") +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"TAvg.pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=PrChangeData, aes(x=Longitude, y=Latitude, fill=Tave), alpha=0.8) + scale_fill_gradient2(midpoint=0, low="#440154FF", mid="#21908CFF", high="#FDE725FF", name="Avg Temp (degC)") + #, limits=c(-100,100)) + 
      ggtitle("Average Temperature (degC)") +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
```

##Update Shap loop
```{r}
rcp45 = T
do_shap = T
#Do_shap F has to be run by hand to generate the predicted effects

  Delta_temp_levels = c(2,2)
  CO2.ppm_levels = c(116,116)
  Adaptation_levels = c("No","Yes") #c(1,2)

#map_df <- NULL
#map_df_all <- NULL`
#map_df_all_crop <- NULL

start <- Sys.time()
numbers <- c(1,2,3,4)
for(k in numbers){ 
  crop_species_map = crop_species[[k]]
  #Filter by cropping areas
  Area_crop<-read.csv(paste0(mdatadir,crop_species_map,"_country.csv"), header=T)
  names(Area_crop)<-c("ID","Longitude","Latitude","IRC","RFC","Total","Country","Flag")
  TAB_in<-right_join(Area_crop,InputData, by=c("Latitude","Longitude"))
  TAB_in<-TAB_in[TAB_in$Total*TAB_in$Flag>0,]
  
    mod.rf.map <- save.mod.rf[[1]][[k]]
    Training_map <- save.Training.rf[[1]][[k]]

    for(levelindex in 1:length(Delta_temp_levels)){
      #map_df[[levelindex]] <- 
      do_shaps(Delta_temp_levels[levelindex], CO2.ppm_levels[levelindex], Adaptation_levels[levelindex], mod.rf.map, do_shap, TAB_in, levelindex, k)
    }
   
    #map_df_all[[1]] <- do.call(rbind.data.frame,map_df)

    #map_df_all_crop[[k]] <- map_df_all[[1]]
  
}
end <- Sys.time()
save.time <- end - start
```

###Plot Shapley Maps
```{r}
shap_df <- NULL
gg1 <- makebaseplot()
cbPalette <- c("#000000","#E69F00", "#009E73", "#0072B2", "#CC79A7", "#D55E00", "#56B4E9", "#F0E442")

for(AdaptLevel in 1:2){
for(k in 1:length(crop_species)){

load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_co2ppm_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_TempAvg_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_Delta_temp_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_Adaptation_4.5.RData"))
load(paste0(outputdir,crop_species[k], "_Nb1_Adapt",AdaptLevel,"_shap_Precip_4.5.RData"))

#Filter by cropping areas
#TAB_p<-read.csv(paste0(mdatadir,"tave_201120.csv"), header=T)
#TAB_p<-na.omit(TAB_p)
  Area_crop<-read.csv(paste0(mdatadir,crop_species[[k]],"_country.csv"), header=T)
  names(Area_crop)<-c("ID","Longitude","Latitude","IRC","RFC","Total","Country","Flag")
  TAB_in<-right_join(Area_crop,InputData, by=c("Latitude","Longitude"))
  TAB_in<-TAB_in[TAB_in$Total*TAB_in$Flag>0,]
  #N<-nrow(TAB_in)

      map_df <- map_of_scenario(AdaptLevel)

      shap_df[[k]] <- map_df %>% tidyr::pivot_wider(names_from="Variable",values_from="Effect")
      
    ## Rename levels (I used data.table::melt, so they were characters)
map_df$Variable <- factor(map_df$Variable, 
                        levels=c("Adaptation", "co2ppm", "Delta_temp", "TempAvg", "Precip"),
                        labels=c("Adaptation", expression(Delta~"CO"[2]~"ppm above 390 ppm"), expression(Delta~" Temperature"~degree~"C"), expression("Average Temperature"~degree~"C"), expression("AveragePrecipitation(mm)")))

    pdf(file=paste0(figoutdir,"shap_rcp45_",crop_species[k],"_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=map_df, aes(x=x, y=y, fill=Effect), alpha=0.8) + scale_fill_gradient2(midpoint=0, low="#440154FF", mid="#21908CFF", high="#FDE725FF", name="Shapley Value") + #, limits=c(-100,100)) + 
      ggtitle(paste0(crop_species[k])) +
      facet_wrap(Variable ~ ., nrow=2, labeller = label_parsed) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()
    
#Max abs Shap value plots
absshaps <- cbind(abs(shap_TempAvg), abs(shap_Delta_temp), abs(shap_co2ppm), abs(shap_Adaptation), abs(shap_Precip))
shaps <- cbind(shap_TempAvg, shap_Delta_temp, shap_co2ppm, shap_Adaptation, shap_Precip)
colnames(absshaps) <- c("Temperature","DeltaTemp","CO2ppm","Adaptation","Precipitation")
colnames(shaps) <- c("Temperature","DeltaTemp","CO2ppm","Adaptation","Precipitation")

shap_df[[k]]$maxabsshaps <- colnames(absshaps)[apply(absshaps,1,which.max)]
shap_df[[k]]$maxshaps <- colnames(shaps)[apply(shaps,1,which.max)]
shap_df[[k]]$minshaps <- colnames(shaps)[apply(shaps,1,which.min)]
shap_df[[k]]$Crop <- rep(crop_species[k], dim(shap_df[[k]])[1])

pdf(file=paste0(figoutdir,"shaphist_",crop_species[k],"_Adapt",AdaptLevel,".pdf"), height=5, width=5)
par(mfrow=c(2,2))
hist(shap_TempAvg, main=expression("Average Temperature"~degree~"C"), xlab="Shapley Value")
hist(shap_Delta_temp, main=expression(Delta~" Temperature"~degree~"C"), xlab="Shapley Value")
hist(shap_co2ppm, main=expression(Delta~"CO"[2]~"ppm above 390 ppm"), xlab="Shapley Value")
hist(shap_Adaptation, main=expression("Adaptation"), xlab="Shapley Value")
hist(shap_Precip, main=expression("Average Precipitation (mm)"), xlab="Shapley Value")
dev.off()

}

shap_df_map <- do.call("rbind", shap_df)
var.labs <- c(expression("Adaptation"), expression(Delta~"CO"[2]~"ppm"), expression(Delta~" Temperature"~degree~"C"), expression("Average Precipitation (mm)"), expression("Average Temperature"~degree~"C"))

    pdf(file=paste0(figoutdir,"shapmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=shap_df_map, aes(x=x, y=y, fill=maxshaps), alpha=0.8) + scale_fill_manual(values = c(cbPalette[5], cbPalette[6], cbPalette[7], cbPalette[4], cbPalette[8]), labels=var.labs) +  
      facet_wrap(Crop ~ ., nrow=2) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_blank()) #change font size of legend title
    )
    dev.off()
    
      pdf(file=paste0(figoutdir,"barplot_shapmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(ggplot(shap_df_map, aes(x=maxshaps)) + geom_bar() + scale_fill_manual(values = c(cbPalette[5], cbPalette[6], cbPalette[7], cbPalette[4], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), 
      axis.text = element_text(size=12),
      axis.text.x=element_text(size=8, angle = 45, vjust = 0.5),
      axis.text.y=element_text(size=8), 
      plot.title=element_text(size=8), 
      legend.text=element_text(size=8), 
      legend.title=element_blank()) +
      ylab("Count") +
      xlab("") +
  scale_x_discrete("", labels = var.labs)
    )
    
        pdf(file=paste0(figoutdir,"shapmin_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=shap_df_map, aes(x=x, y=y, fill=minshaps), alpha=0.8) + scale_fill_manual(values = c(cbPalette[5], cbPalette[6], cbPalette[7], cbPalette[4], cbPalette[8]), labels=var.labs) +
      facet_wrap(Crop ~ ., nrow=2) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_blank()) #change font size of legend title
    )
    dev.off()
    
      pdf(file=paste0(figoutdir,"barplot_shapmin_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(ggplot(shap_df_map, aes(x=minshaps)) + geom_bar() + scale_fill_manual(values = c(cbPalette[5], cbPalette[6], cbPalette[7], cbPalette[4], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), 
      axis.text = element_text(size=12),
      axis.text.x=element_text(size=8, angle = 45, vjust = 0.5),
      axis.text.y=element_text(size=8), 
      plot.title=element_text(size=8), 
      legend.text=element_text(size=8), 
      legend.title=element_blank()) +
      ylab("Count") +
      xlab("") +
  scale_x_discrete("", labels = var.labs)
    )
    
        pdf(file=paste0(figoutdir,"shapabsmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=shap_df_map, aes(x=x, y=y, fill=maxabsshaps), alpha=0.8) + scale_fill_manual(values = c(cbPalette[5], cbPalette[6], cbPalette[7], cbPalette[4], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_blank()) #change font size of legend title
    )
    dev.off()
    
        pdf(file=paste0(figoutdir,"barplot_shapabsmax_Adapt",AdaptLevel,".pdf"), height=6, width=9)
    print(ggplot(shap_df_map, aes(x=maxabsshaps)) + geom_bar() + scale_fill_manual(values = c(cbPalette[5], cbPalette[6], cbPalette[7], cbPalette[4], cbPalette[8]), labels=var.labs) + 
      facet_wrap(Crop ~ ., nrow=2) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), 
      axis.text = element_text(size=12),
      axis.text.x=element_text(size=8, angle = 45, vjust = 0.5),
      axis.text.y=element_text(size=8), 
      plot.title=element_text(size=8), 
      legend.text=element_text(size=8), 
      legend.title=element_blank()) +
      ylab("Count") +
      xlab("") +
  scale_x_discrete("", labels = var.labs)
    )
    dev.off()
    
save(shap_df_map, file=paste0(outputdir,AdaptLevel,"_ShapdfMap.RData"))
}
```

#Fig2 Yield Change Summary Map
```{r}
map_df_adapt <- NULL
map_df_noadapt <- NULL
map_df_all <- NULL

#Effect is the only different column - rbind to combine all Nbs, labelling the crop
for (k in 1:4){
  load(file=paste0(outputdir,crop_species[k], "_Nb",1,"_Adapt",1,"_predicted_effect_4.5.RData"))
  
  map_df_noadapt[[k]] <- NewData
  map_df_noadapt[[k]]$Crop <- crop_species[k]
  
  load(file=paste0(outputdir,crop_species[k], "_Nb",1,"_Adapt",2,"_predicted_effect_4.5.RData"))
  
  map_df_adapt[[k]] <- NewData
  map_df_adapt[[k]]$Crop <- crop_species[k]
}

map_df_noadapt[[2]]$Adaptation <- rep("No", dim(map_df_noadapt[[2]])[1])
map_df_noadapt[[4]]$Adaptation <- rep("No", dim(map_df_noadapt[[4]])[1])
map_df_adapt[[2]]$Adaptation <- rep("Yes", dim(map_df_adapt[[2]])[1])
map_df_adapt[[4]]$Adaptation <- rep("Yes", dim(map_df_adapt[[4]])[1])

map_df_all <- map_df_adapt[[1]] %>% full_join(map_df_adapt[[2]]) %>% full_join(map_df_adapt[[3]]) %>% full_join(map_df_adapt[[4]]) %>% full_join(map_df_noadapt[[1]]) %>% full_join(map_df_noadapt[[2]]) %>% full_join(map_df_noadapt[[3]]) %>% full_join(map_df_noadapt[[4]]) 

# #average Nbs for each crop #not used as now only one Nb
# map_summary <- map_combined %>%
#   group_by(x, y, Delta_temp_level, CO2.ppm_level, Adaptation_level, Crop) %>%
#   dplyr::summarise(Mean_Effect = mean(Effect), 
#                    SE_Effect = sd(Effect)/sqrt(n()),
#                    n = n())

# New facet label names for Adaptation variable
adpt.labs <- c("No Adaptation", "Adaptation")
names(adpt.labs) <- c("No", "Yes")

map_df_all$Crop <- as.factor(map_df_all$Crop)
#levels(map_df_all$Crop) <- c("Maize", "Rice", "Soybean", "Wheat")

###Plot####
    pdf(file=paste0(figoutdir,"allcrops_rcp45",rcp45,".pdf"), height=6, width=9)
    print(gg1 + geom_tile(data=map_df_all, aes(x=Longitude, y=Latitude, fill=Effect), alpha=0.8) + scale_fill_gradient2(midpoint=0, low="#440154FF", mid="#21908CFF", high="#FDE725FF", limits=c(-100,100), name="Yield Change (%)") + 
      ggtitle(paste0("RCP 4.5")) +
      facet_grid(Adaptation ~ Crop, labeller = labeller(Adaptation = adpt.labs)) +
      theme(#text=element_text(size=8), #change font size of all text
        axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
        axis.text.y=element_text(size=8), #change font size of axis titles
        plot.title=element_text(size=8), #change font size of plot title
        legend.text=element_text(size=8), #change font size of legend text
        legend.title=element_text(size=8)) #change font size of legend title
    )
    dev.off()

      pdf(file=paste0(figoutdir,"allcrops_rcp45",rcp45,"_boxplot.pdf"), height=3, width=5)
          print(ggplot(data=map_df_all, aes(x=Crop, y=Effect, fill=Adaptation), alpha=0.8) + 
                  geom_boxplot() + stat_boxplot(geom ='errorbar') +
                  ggtitle(paste0("RCP 4.5")) + theme_classic() +
                  geom_hline(yintercept=0, linetype='dotted') +
          # theme(#text=element_text(size=8), #change font size of all text
          #   axis.text.x=element_text(size=8, angle = 45, vjust = 0.5), #change font size of axis text
          #   axis.text.y=element_text(size=8), #change font size of axis titles
          #   plot.title=element_text(size=8), #change font size of plot title
          #   legend.text=element_text(size=8), #change font size of legend text
          #   legend.title=element_text(size=8)) +
              ylab("Yield Change (%)") #change font size of legend title
          )
      dev.off()
```


# Shap Value Boxplots
```{r}
#WORKING - needs to be outside crop loop
# shap_df_all <- map_df %>% pivot_longer()
# 
#     pdf(file=paste0(figoutdir,"shap_box_",crop_species[k],"_Adapt",AdaptLevel,".pdf"), height=6, width=9)
#               print(ggplot(data=map_df_all, aes(x=Crop, y=Effect, fill=Adaptation), alpha=0.8) + 
#                   geom_boxplot() + stat_boxplot(geom ='errorbar') +
#                   ggtitle(paste0("RCP 4.5")) + theme_classic() +
#                   geom_hline(yintercept=0, linetype='dotted') +
#               ylab("Shapley Value") #change font size of legend title
#           )
#     print(boxplot geom_tile(data=map_df, aes(x=x, y=y, fill=Effect), alpha=0.8) +
#       ggtitle(paste0(crop_species[k])) +
#       facet_wrap(Variable ~ ., nrow=2, labeller = label_parsed)
#     )
#     dev.off()
```

# Formatted Fig3plot
## Load shapmaps and summaries
```{r}
load(paste0(outputdir,"1_ShapdfMap.RData"))
adpt1shap <- shap_df_map
load(paste0(outputdir,"2_ShapdfMap.RData"))
adpt2shap <- shap_df_map

#Summaries
table(adpt1shap$maxabsshaps, by=adpt1shap$Crop)
table(adpt1shap$maxshaps, by=adpt1shap$Crop)
table(adpt1shap$minshaps, by=adpt1shap$Crop)

table(adpt2shap$maxabsshaps, by=adpt2shap$Crop)
table(adpt2shap$maxshaps, by=adpt2shap$Crop)
table(adpt2shap$minshaps, by=adpt2shap$Crop)

summary(adpt1shap[adpt1shap$Crop=="Maize",]$Adaptation)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$Adaptation)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$TempAvg)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$TempAvg)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$Precip)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$Precip)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$Delta_temp)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$Delta_temp)
summary(adpt1shap[adpt1shap$Crop=="Maize",]$co2ppm)
summary(adpt2shap[adpt2shap$Crop=="Maize",]$co2ppm)


summary(adpt1shap[adpt1shap$Crop=="Rice",]$Adaptation)
summary(adpt2shap[adpt2shap$Crop=="Rice",]$Adaptation)
summary(adpt1shap[adpt1shap$Crop=="Rice",]$TempAvg)
summary(adpt2shap[adpt2shap$Crop=="Rice",]$TempAvg)

summary(adpt1shap[adpt1shap$Crop=="Soybean",]$Adaptation)
summary(adpt2shap[adpt2shap$Crop=="Soybean",]$Adaptation)
summary(adpt1shap[adpt1shap$Crop=="Soybean",]$TempAvg)
summary(adpt2shap[adpt2shap$Crop=="Soybean",]$TempAvg)
```

##Merge and plot
```{r}
adptshap <- rbind(adpt1shap, adpt2shap)

#Checking if all same - Rice is maybe okay but Soybean is the same :(
all.equal(adpt1shap[adpt1shap$Crop=="Maize",], adpt2shap[adpt2shap$Crop=="Maize",])
all.equal(adpt1shap[adpt1shap$Crop=="Rice",], adpt2shap[adpt2shap$Crop=="Rice",])
all.equal(adpt1shap[adpt1shap$Crop=="Wheat",], adpt2shap[adpt2shap$Crop=="Wheat",])
all.equal(adpt1shap[adpt1shap$Crop=="Soybean",], adpt2shap[adpt2shap$Crop=="Soybean",])

plot(adpt1shap[adpt1shap$Crop=="Maize",]$Precip, adpt2shap[adpt2shap$Crop=="Maize",]$Precip)
plot(adpt1shap[adpt1shap$Crop=="Rice",]$Precip, adpt2shap[adpt2shap$Crop=="Rice",]$Precip)
plot(adpt1shap[adpt1shap$Crop=="Wheat",]$Precip, adpt2shap[adpt2shap$Crop=="Wheat",]$Precip)
plot(adpt1shap[adpt1shap$Crop=="Soybean",]$Precip, adpt2shap[adpt2shap$Crop=="Soybean",]$Precip)
```

#Summaries
```{r}
######Summary stats#####
summary_stats <- map_df_all %>%
  group_by(Delta_temp, CO2.ppm, Adaptation, Crop) %>%
  dplyr::summarise(Mean = mean(Effect), 
                   Median = median(Effect), 
                   SE = sd(Effect)/sqrt(n()),
                   min = min(Effect),
                   max = max(Effect),
                   n = n())
summary_stats
```


